---
title: "Spatial Covariates"
output: pdf_document
date: "2023-04-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The understory fuels generation algorithm.

1. Draw a number of fuel elements $n \sim Pois(\lambda*X*Y)$ where $\lambda$ is the relative density parameter for a domain $[0,X]\times[0,Y]$

2. Randomly sample $x,y$ locations $[\boldsymbol{x}|\boldsymbol{y}]\sim Unif([0,X]\times[0,Y])$

```{r, echo=F}
dimX = 30;dimY = 30
lambda = 1
n = rpois(1,lambda*dimX*dimY)
X = runif(n,0,dimX); Y = runif(n,0,dimY)
df = data.frame(x=X,y=Y)
ggplot2::ggplot() + ggplot2::geom_point(data=df,ggplot2::aes(x,y))
```

3. Sample from a mean-zero intensity function (Gaussian Process) with heterogeneity parameter $\rho$ and variance $\sigma^2$ at these locations $\tilde{G}(\boldsymbol{x},\boldsymbol{y}|\rho,\sigma)$.

```{r, echo=F}
mod = fuelsgen:::GP_fit(xy=cbind(X,Y),rho=5)
GP = fuelsgen:::GP_sample(mod,sd=3)
```

4. Determine the probability of keeping a shrub at each location $\boldsymbol{p}=plogis(\tilde{G}(\boldsymbol{x},\boldsymbol{y}|\rho,\sigma))$

```{r, echo=F}
df$p = plogis(GP)
ggplot2::ggplot() + ggplot2::geom_point(data=df,ggplot2::aes(x,y,col=p)) + ggplot2::scale_color_continuous(limits=c(0,1))
```

5. Keep shrubs at locations where $\boldsymbol{u}<\boldsymbol{p}$ $(\boldsymbol{u} \sim Unif(0,1))$

```{r, echo=F}
df$keep = as.factor(runif(n) < df$p)
ggplot2::ggplot() + ggplot2::geom_point(data=df,ggplot2::aes(x,y,col=keep))
```

# Incorporate spatial covariates

If we wish to include spatial covariates $\boldsymbol{X}$ with effect $\boldsymbol{\beta}$ we add them to the GP realization before downsampling with the logistic.

4. Determine the probability of keeping a shrub at each location $\boldsymbol{p}=plogis(\tilde{G}(\boldsymbol{x},\boldsymbol{y}|\rho,\sigma) + \boldsymbol{X}\boldsymbol{\beta})$\newline

At this moment, the code is technically usable with any spatial covariate, however we have designed it so that $X$ is scaled between $[-1,1]$ and $\beta \in \mathbb{R^+}$. This is so a single X can have both a positive and negative effect of shrub probability. If $X \in \mathbb{R}^+$ then X will always have a positive effect, which may not be the desired effect, especially since a mean zero Intensity function gives an average probability of $50\%$ to each location, and we may want to significantly reduce that probability with a covariate.

## How to choose $\sigma^2$ and $\boldsymbol{\beta}$?

This plot may be useful in determining a balance between the GP standard deviation (heterogeneity scale) and the spatial covariate parameters $\boldsymbol{\beta}$. Below we show the probability of fuel placement based on the value of $GP+X\boldsymbol{\beta}$ at any given location. We can see that at $-5,5$ the probability of fuel placement is less than $1\%$ and greater than $99\%$ respectively. Desired trade-offs between the two terms may be achievable with $heterogeneity\;scale \in [0,5/3]$ and $\boldsymbol{\beta} \leq 5$. These are reasonable starting points to consider for these parameters but there may be circumstances where values of greater magnitude help to achieve a desired effect in fuel placement.

```{r, fig.width=10,fig.height=6, echo=F}
x = seq(-10,10,.1)
plot(x,plogis(x),type='l',ylab='fuel probability',xlab='GP+XB')
axis(1,at=seq(-10,10,1))
abline(v=qlogis(c(.01,.05,.95,.99)),lty=1:4)
legend('topleft',inset=c(.05,.05),legend=c('1%','5%','95%','99%'),lty=1:4)
```
