---
title: "Bandelier Calibration"
output: pdf_document
date: "2024-03-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(terra)
library(here)
library(ggplot2)
library(dplyr)
library(fuelsgen)
source('~/Desktop/R tools/unit_xform.R')
source('~/Desktop/R tools/Metro_Hastings_Stochastic.R')
```

Do clustering of scan and grid locations

```{r, echo=F}
load(here('examples/bandelier/results/topology_df.RData'))
load(here('examples/bandelier/results/plot_nshrub.RData'))
plot_topo$Plot = NULL
nplot = nrow(plot_topo)

all_topo = as.data.frame(unit_xform(rbind(plot_topo,grid_topo)))
all_topo$coverage = c(plot_topo$coverage,grid_topo$coverage)
all_topo$type = c(rep('plot',nrow(plot_topo)),rep('grid',nrow(grid_topo)))

ggplot(all_topo) + geom_point(aes(x_crd,y_crd,col=coverage,shape=type)) + scale_color_viridis_c()
hist(plot_topo$coverage)
hist(grid_topo$coverage)

my_dist = function(X,w=rep(1,ncol(X))){
  # slow and expensive, but it doesn't really matter for this application
  names = colnames(X)
  X = t(t(X) * sqrt(w))
  if(!('aspect' %in% names)){
    return(dist(X))
  } else{
    d = as.matrix(dist(X))
    which_aspect = which(names=='aspect')
    aspect_d = as.matrix(dist(X[,which_aspect,drop=F]))
    fix = which(aspect_d>.5) # fix any aspects that are greater than .5 apart
    d[fix] = sqrt(d[fix]^2-aspect_d[fix]^2) # remove the aspect part of the distance calculation from these points
    aspect_d = pmin(aspect_d,1-aspect_d)
    d[fix] = sqrt(d[fix]^2 + aspect_d[fix]^2)
    return(as.dist(d))
  }
}

# search all possible combinations of these 6 parameters for clustering
which.X.cl = list()
vars = c('slope','aspect','elevation','coverage','x_crd','y_crd')
for(i in 1:length(vars)){
  which.X.cl[[i]] = t(combn(vars,i))
}
for(i in 1:length(which.X.cl)){
  for(j in 1:nrow(which.X.cl[[i]])){
    id = which.X.cl[[i]][j,]
    w = numeric(length(id))+1
    for(k in 1:length(id)){
      if(id[k]%in%c('x_crd','y_crd')){
        w[k] = .05
      }
    }
    d = my_dist(all_topo[,id],w)
    cl.model = hclust(d)
    cl = cutree(cl.model,3)
    all_topo$cl = cl
    cl.dens = data.frame(d=nshrub,clust=cl[1:nplot])
    print(ggplot(cl.dens) + geom_point(aes(as.factor(clust),d)) + labs(title=paste(which.X.cl[[i]][j,],collapse = ', ')))
  }
}

# I like slope, aspect, coverage when coverage is available
which.X.cl.naip = c('elevation','coverage')
d.naip = my_dist(all_topo[,which.X.cl.naip])

cl.model.naip = hclust(d.naip)
cl.naip = cutree(cl.model.naip,3)
all_topo$cl.naip = cl.naip
cl.dens.naip = data.frame(d=nshrub,clust=cl.naip[1:nplot])
print(ggplot(cl.dens.naip) + geom_point(aes(as.factor(clust),d)) + labs(title=paste(which.X.cl.naip,collapse = ', ')))
    
# When coverage is not available, I like slope and x_crd or y_crd, they are both about the same
which.X.cl = c('aspect')
d = my_dist(all_topo[,which.X.cl],w = c(1))

cl.model = hclust(d)
cl = cutree(cl.model,3)
all_topo$cl = cl
cl.dens = data.frame(d=nshrub,clust=cl[1:nplot])
print(ggplot(cl.dens) + geom_point(aes(as.factor(clust),d)) + labs(title=paste(which.X.cl,collapse = ', ')))
```

```{r vis clusters, echo=F}
# NO NAIP
ggplot(all_topo) + geom_point(aes(x=x_crd,y=y_crd,col=as.factor(cl),pch=as.factor(type)))
ggplot(subset(all_topo,type=='plot')) + geom_point(aes(x=x_crd,y=y_crd,col=as.factor(cl)))
ggplot(cl.dens) + geom_boxplot(aes(as.factor(clust),d))

# NAIP
ggplot(all_topo) + geom_point(aes(x=x_crd,y=y_crd,col=as.factor(cl.naip),pch=as.factor(type)))
ggplot(subset(all_topo,type=='plot')) + geom_point(aes(x=x_crd,y=y_crd,col=as.factor(cl.naip)))
ggplot(cl.dens.naip) + geom_point(aes(as.factor(clust),d))
```
Put fuel objects into clusters for calibration

```{r, echo=FALSE}
load(here('examples/bandelier/results/plot_data.RData'))
ncl = 3
data.cl = vector(mode='list',length=ncl)
data.cl.naip = vector(mode='list',length=ncl)
cl.plot = cl[1:nplot]
cl.plot.naip = cl.naip[1:nplot]
for(i in 1:ncl){
  data.cl[[i]] = gen_fuels(dimX = 20, dimY = 20,
                           density = .05,        # shrubs per unit area
                           heterogeneity = 3,     # level of heterogeneity in shrub placement
                           heterogeneity.scale = 1,      # scale of the mean-zero GP realization
                           radius = 1, sd_radius = .1,  # normal distribution parameters for shrub radius
                           height = NULL, sd_height = 0, # normal distribution parameters for shrub height
                           reps=1,                      # number of random maps to generate
                           GP.init.size=32,             # How densely to sample GP (n^3 scaling, <=1000 is pretty fast)
                           seed=10,                      # random seed for reproducibility
                           parallel=F,verbose = F)
  data.cl[[i]]$dat = data$dat[cl.plot==i]
  data.cl[[i]]$reps = length(data.cl[[i]]$dat)
  data.cl[[i]]$mets = mets
  data.cl[[i]]$mets$mets = data.cl[[i]]$mets$mets[cl.plot==i,]
  data.cl[[i]]$nshrub = numeric(data.cl[[i]]$reps)
  for(k in 1:data.cl[[i]]$reps){
    data.cl[[i]]$nshrub[k] = nrow(data.cl[[i]]$dat[[k]])
  }
  data.cl.naip[[i]] = gen_fuels(dimX = 20, dimY = 20,
                                 density = .05,        # shrubs per unit area
                                 heterogeneity = 3,     # level of heterogeneity in shrub placement
                                 heterogeneity.scale = 1,      # scale of the mean-zero GP realization
                                 radius = 1, sd_radius = .1,  # normal distribution parameters for shrub radius
                                 height = NULL, sd_height = 0, # normal distribution parameters for shrub height
                                 reps=1,                      # number of random maps to generate
                                 GP.init.size=100,             # How densely to sample GP (n^3 scaling, <=1000 is pretty fast)
                                 seed=10,                      # random seed for reproducibility
                                 parallel=F,verbose = F)
    
  data.cl.naip[[i]]$dat = data$dat[cl.plot.naip==i]
  data.cl.naip[[i]]$reps = length(data.cl.naip[[i]]$dat)
  data.cl.naip[[i]]$nshrub = numeric(data.cl.naip[[i]]$reps)
  for(k in 1:data.cl.naip[[i]]$reps){
    data.cl.naip[[i]]$nshrub[k] = nrow(data.cl.naip[[i]]$dat[[k]])
  }
  data.cl.naip[[i]]$mets = mets
  data.cl.naip[[i]]$mets$mets = data.cl.naip[[i]]$mets$mets[cl.plot.naip==i,]
}
```

# Naip Calibration

```{r, echo=F}
prior.naip = vector(mode='list',length = ncl)
mcmc.naip = vector(mode='list',length = ncl)
for(i in 2:ncl){
  if(data.cl.naip[[i]]$reps<10){
    # not enough observations to estimate the covariance matrix augment with simulations
    prior.naip[[i]] = get_prior_info(data.cl.naip[[i]],data.cl.naip[[i]]$mets,est_cov_obs = F)
  } else{
    prior.naip[[i]] = get_prior_info(data.cl.naip[[i]],data.cl.naip[[i]]$mets)
  }
  mcmc.naip[[i]] = mcmc_MH_adaptive(data.cl.naip[[i]]$mets,
                                    data.cl.naip[[i]],
                                    prior.naip[[i]],
                                    paste0(here('examples/bandelier/results/mcmc/calib_naip_cl_'),i,'.RData'),
                                    adapt.par = c(100,20,.5,.5),
                                    n.samples=10000,n.burn=0,gen_reps=25,avg='mets',
                                    gen_parallel=F,mets_parallel=T)
}
```

# No Naip calibration

```{r, echo=F}
prior = vector(mode='list',length = ncl)
mcmc = vector(mode='list',length = ncl)
for(i in 1:ncl){
  if(data.cl[[i]]$reps<10){
    # not enough observations to estimate the covariance matrix with obs alone, augment with simulations
    prior[[i]] = get_prior_info(data.cl[[i]],data.cl[[i]]$mets,est_cov_obs = F)
  } else{
    prior[[i]] = get_prior_info(data.cl[[i]],data.cl[[i]]$mets)
  }
  mcmc[[i]] = mcmc_MH_adaptive(data.cl[[i]]$mets,
                                    data.cl[[i]],
                                    prior[[i]],
                                    paste0(here('examples/bandelier/results/mcmc/calib_cl_'),i,'.RData'),
                                    adapt.par = c(100,20,.5,.5),
                                    n.samples=10000,n.burn=0,gen_reps=25,avg='mets',
                                    gen_parallel=F,mets_parallel=T)
}
```

# naip calibration results

```{r, echo=F}
# what are the mle distributions, are they the same as the bayesian ones i have?
par(mfrow=c(2,2))
ncl = 3
for(j in 1:ncl){
  load(paste0(here('examples/bandelier/results/mcmc/calib_naip_cl_'),j,'.RData'))
  for(i in 1:4){
    plot(density(mcmc$trace[(.5*10000):10000,i],
                 from=mcmc$prior$prior_params$lb[i],
                 to=mcmc$prior$prior_params$ub[i]),
         main='',xlab=mcmc$par_names[i])
    abline(v=mcmc$prior$theta_est[i],col='red')
  }
  for(i in 1:4){
    plot(mcmc$trace[(.5*10000):10000,i],type='l',
         main='',xlab=mcmc$par_names[i],ylab='')
    abline(h=mcmc$prior$theta_est[i],col='red')
  }
}
j=1;i=4
load(paste0(here('examples/bandelier/results/mcmc/calib_naip_cl_'),j,'.RData'))
plot(density(mcmc$trace[(.5*10000):10000,i],from=0),xlim=c(0,.1),ylim=c(0,350),
        main='',xlab=mcmc$par_names[i],col='red');abline(v=mcmc$prior$theta_est[i],col='red')
j=2;
load(paste0(here('examples/bandelier/results/mcmc/calib_naip_cl_'),j,'.RData'))
lines(density(mcmc$trace[(.5*10000):10000,i],from=0),col='green3',
        main='',xlab=mcmc$par_names[i]);abline(v=mcmc$prior$theta_est[i],col='green3')
j=3;
load(paste0(here('examples/bandelier/results/mcmc/calib_naip_cl_'),j,'.RData'))
lines(density(mcmc$trace[(.5*10000):10000,i],from=0),col='blue',
        main='',xlab=mcmc$par_names[i]);abline(v=mcmc$prior$theta_est[i],col='blue')
```

# no naip calibration results

```{r, echo=F}
par(mfrow=c(2,2))
ncl = 3
for(j in 1:ncl){
  load(paste0(here('examples/bandelier/results/mcmc/calib_cl_'),j,'.RData'))
  for(i in 1:4){
    plot(density(mcmc$trace[(.75*5000):5000,i],from=mcmc$prior$prior_params$lb[i],to=mcmc$prior$prior_params$ub[i]),
         main='',xlab=mcmc$par_names[i])
    abline(v=mcmc$prior$theta_est[i],col='red')
  }
  # for(i in 1:4){
  #   plot(mcmc$trace[(.75*5000):5000,i],type='l',
  #        main='',xlab=mcmc$par_names[i])
  #   abline(h=mcmc$prior$theta_est[i],col='red')
  # }
}
par(mfrow=c(1,1))
j=1;i=4
load(paste0(here('examples/bandelier/results/mcmc/calib_cl_'),j,'.RData'))
plot(density(mcmc$trace[(.75*5000):5000,i],from=0),xlim=c(0,.1),ylim=c(0,350),
        main='',xlab=mcmc$par_names[i],col='red');abline(v=mcmc$prior$theta_est[i],col='red')
# hist(data.cl[[j]]$nshrub,freq=F,add=T)
j=2;
load(paste0(here('examples/bandelier/results/mcmc/calib_cl_'),j,'.RData'))
lines(density(mcmc$trace[(.75*5000):5000,i],from=0),col='green2',
        main='',xlab=mcmc$par_names[i]);abline(v=mcmc$prior$theta_est[i],col='green2')
j=3;
load(paste0(here('examples/bandelier/results/mcmc/calib_cl_'),j,'.RData'))
lines(density(mcmc$trace[(.75*5000):5000,i],from=0),col='blue',
        main='',xlab=mcmc$par_names[i]);abline(v=mcmc$prior$theta_est[i],col='blue')
legend('topright',legend=c('1','2','3'),col=c('red','green2','blue'),lty=1)
```

# generate naip fuels

```{r, echo=F}
reps = 10
params = matrix(nrow=reps,ncol=4)
fuels = vector(mode='list',length=nrow(all_topo))
for(j in 1:nrow(all_topo)){
  if(j%%10==0){cat(j,' ')}
  fuels[[j]] = vector(mode='list',length=reps)
  # for each grid cell generate fuels w.r.t. cluster membership
  cluster = all_topo$cl.naip[j]
  load(paste0(here('examples/bandelier/results/mcmc/calib_naip_cl_'),cluster,'.RData'))
  params = mcmc$trace[sample(5001:nrow(mcmc$trace),10,F),]
  for(i in 1:reps){
     fuels[[j]][[i]] = gen_fuels(20,20,params[i,4],params[i,2],params[i,3],1,0,params[i,1],GP.init.size = 32,verbose = F)
  }
}
save(fuels,file='~/fuelsgen/examples/bandelier/generated_fuels/naip/naip_fuelsgen_object.RData')

# linear model from radius to height
r = h = numeric()
for(i in 1:length(data$dat)){
  r = c(r,data$dat[[i]]$r)
  h = c(h,data$dat[[i]]$h)
}
df_rh = data.frame(r=r,h=h)
h3 = which(h>2.99)
df_rh$h3 = 0
df_rh$h3[h3] = 1
mod_rh = lm(h~1+r,df_rh) # model using all data
mod_rh_3 = lm(h~1+r,subset(df_rh,h3==0)) # model not including data where heights have been truncated

plot(df_rh$r,df_rh$h,xlim=c(0,3),ylim=c(0,3))
lines(df_rh$r,mod_rh$fitted.values,col='red')
lines(df_rh$r[df_rh$h3==0],mod_rh_3$fitted.values,col='blue')

# convert to 10 data frames and sample heights
# niko has 1040 and 980
Xmax = 1020 # (coords to 1020)
Ymax = 960 # coords to 960
dfs = vector(mode='list',length=10)
for(i in 1:10){
  for(j in (nplot+1):nrow(all_topo)){ # accidentally generated fuels for plot [1:155] locations, only need grid locations [156:2548]
    dfs[[i]] = rbind(dfs[[i]],data.frame(X=fuels[[j]][[i]]$dat[[1]]$X + (all_topo$x_crd[j] * Xmax),
                                         Y=fuels[[j]][[i]]$dat[[1]]$Y + (all_topo$y_crd[j] * Ymax),
                                         r=fuels[[j]][[i]]$dat[[1]]$r))
  }
  # sample from truncated normal for heights
  rs = predict(mod_rh,data.frame(r=dfs[[i]]$r))
  for(k in 1:length(rs)){
    dfs[[i]]$h[k] = truncdist::rtrunc(1,'norm',a=1,b=3,mean=rs[k],sd=sigma(mod_rh))
  }
}

pdf(paste0('~/fuelsgen/examples/bandelier/generated_fuels/naip/gen_','all','.pdf'),width = 10,height = 9.4)
par(mfrow=c(3,4),mar=c(1,1,1,1))
for(i in 1:10){
  plot(dfs[[i]]$X,dfs[[i]]$Y,pch=16,cex=.2,asp=Ymax/Xmax)
}
dev.off()
for(i in 1:10){
  write.csv(dfs[[i]],file = paste0('~/fuelsgen/examples/bandelier/generated_fuels/naip/gen_',i,'.csv'))
}
```

# generate no naip fuels

No naip fuels have a total fuel load of about 15000 fuels over the entire domain. By contrast naip fuels have about 23000. This is because the clustering algorithm can't do a great job at distinguishing the highly dense regions without the shrub coverage metric. In this case the no naip clusting uses only aspect. The naip fuels on the other hand have regions of significantly increased density due to the fact that cluster 1 contains only a handful of scans, all with relatively high fuel density.

```{r, echo=F}
reps = 10
params = matrix(nrow=reps,ncol=4)
fuels = vector(mode='list',length=nrow(all_topo))
for(j in 1:nrow(all_topo)){
  if(j%%10==0){cat(j,' ')}
  fuels[[j]] = vector(mode='list',length=reps)
  # for each grid cell generate fuels w.r.t. cluster membership
  cluster = all_topo$cl[j]
  load(paste0(here('examples/bandelier/results/mcmc/calib_cl_'),cluster,'.RData'))
  params = mcmc$trace[sample(5001:nrow(mcmc$trace),10,F),]
  for(i in 1:reps){
     fuels[[j]][[i]] = gen_fuels(20,20,params[i,4],params[i,2],params[i,3],1,0,params[i,1],GP.init.size = 32,verbose = F)
  }
}
save(fuels,file='~/fuelsgen/examples/bandelier/generated_fuels/no naip/fuelsgen_object.RData')

dfs = vector(mode='list',length=10)
for(i in 1:10){
  for(j in (nplot+1):nrow(all_topo)){ # accidentally generated fuels for plot [1:155] locations, only need grid locations [156:2603]
    dfs[[i]] = rbind(dfs[[i]],data.frame(X=fuels[[j]][[i]]$dat[[1]]$X + (all_topo$x_crd[j] * Xmax),
                                         Y=fuels[[j]][[i]]$dat[[1]]$Y + (all_topo$y_crd[j] * Ymax),
                                         r=fuels[[j]][[i]]$dat[[1]]$r))
  }
  # sample from truncated normal for heights
  rs = predict(mod_rh,data.frame(r=dfs[[i]]$r))
  for(k in 1:length(rs)){
    dfs[[i]]$h[k] = truncdist::rtrunc(1,'norm',a=1,b=3,mean=rs[k],sd=sigma(mod_rh))
  }
}
pdf(paste0('~/fuelsgen/examples/bandelier/generated_fuels/no naip/gen_','all','.pdf'),width = 10,height = 9.4)
par(mfrow=c(3,4),mar=c(1,1,1,1))
for(i in 1:10){
  plot(dfs[[i]]$X,dfs[[i]]$Y,pch=16,cex=.2)
}
dev.off()
for(i in 1:10){
  write.csv(dfs[[i]],file = paste0('~/fuelsgen/examples/bandelier/generated_fuels/no naip/gen_',i,'.csv'))
}
```

