---
title: "chm_example"
output: pdf_document
date: "2023-04-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(fuelsgen)
```

In this example we will read in a .tif file of a Canopy Height Model and use it to influence understory fuel generation on its domain.

```{r plot chm, warning=F}
# Canopy
chm = raster::raster(here::here("examples/CHM/LA_CHM.tif"))
raster::plot(chm)

# Road
chm_road = raster::raster(here::here("examples/CHM/LA_CHM_road.tif"))
raster::extent(chm_road) = raster::extent(chm)
raster::values(chm_road)[raster::values(chm_road) == 0] = -1
raster::values(chm_road)[raster::values(chm_road) != -1] = 0
```

In its raw form, the Canopy Height Model raster is not compatible with the understory fuels generation model. There is a helper function that will update the coordinate system of the raster to work with out model. The X and Y coordinates have been changed so that the origin is at (0,0) and the response values have been scaled to [-1,1].

```{r update raster to work with fuelsgen}
chm = modify_raster(chm,type='regular')
chm_road = modify_raster(chm_road,type='zero')

raster::plot(chm$raster)
raster::plot(chm_road$raster,col=colorRampPalette(c("black","white"))(11))
```

The chm object also contains coordinates and values in the correct format for the fuelsgen model.

```{r build fuel map using this chm}
B = c(1,10)
fuel = gen_fuels(chm$dimX, chm$dimY,
                density = .025,           # shrubs per unit area before downsampling
                heterogeneity = 5,        # level of heterogeneity in shrub placement
                heterogeneity.scale = 1,  # scale of the mean-zero GP realization
                radius = 1, sd_radius = 0,# normal distribution parameters for shrub radius
                height = NULL, sd_height = 0,# normal distribution parameters for shrub height
                sp.cov.locs = list(chm$locs[[1]],chm_road$locs[[1]]),
                                          # locations where canopy height is known
                sp.cov.vals = list(chm$vals[[1]],chm_road$vals[[1]]),     
                                          # canopy height values
                sp.cov.scale = B,         # Importance of chm for shrub placement
                reps=1,                   # number of random maps to generate
                GP.init.size=100,         # How densely to sample GP
                seed=10,                  # random seed for reproducibility)
                verbose = F)                 
```

Plot these points over the raster to visualize the effect of the CHM. We can see that the shrub locations are heavily influenced by the CHM. This is because $B=5$ is large compared to the standard deviation of the GP (1). We can imagine creating another spatial covariate representing locations where shrub placement should be probability 0, such as along the road.

```{r raster plot}
raster::plot(chm$raster,legend=F)
points(fuel$dat[[1]]$X,fuel$dat[[1]]$Y,cex=.5,pch=16)
```
